---
title: "Mechanisms of Injury Meta-analysis"
format: html
editor: visual
---

## Using a meta-regression to determine whether TBI etiology produces distinct spectral power signatures 

```{r}
1 + 1
```

Necessary libraries

```{r}
library(metafor)
library(tidyverse)
library(readxl)
library(psych)
library(ggplot2)
library(dplyr)
library(sjPlot)
library(lmerTest)
library(lme4)
```

## Set directory and import data

Imported data frame detailing study, age, sex, severity, mechanism and all 5 bands effect sized and variance.

```{r}
setwd('/Users/nwakamma.m/Documents/manscripts/EEG & TBI review/meta_analysis')
moi_meta <- read_excel('eeg_tbi_meta_analysis.xlsx')
```

## Convert Cohen's D to Hedges G.

Very necessary to adjust for sample size when calculating effect size. Hedge's G takes sample size into account. The escalc package can get hedges G. Used variance estimation formula using hedge's g effect size.

```{r}
g <- esc::hedges_g(0.62, 63)   #input cohen's d and total sample size to get hedge's g

v <- (34 + 29)/(34 * 29) + g^2/(2*(34 + 29 - 2)) #get variance

g

v
```

## Multilevel Meta-regression with covariates

Get study labels

```{r}
moi_meta$mechanism <- as.factor(moi_meta$mechanism)
moi_meta$severity <- as.numeric(moi_meta$severity)
moi_meta$electrode_site <- as.numeric(moi_meta$electrode_site)

study_labels <- c("Thompson et al., 2005 - S",
"Balkan et al., 2015 - S",
"Munia et al., 2017 - S",
"Coyle et al., 2023 - NS",
"Fisher et al., 2015 - NS",
"Korn et al., 2005 - NS",
"Tebano et al., 1988 - NS",
"Williams et al., 2008 - M",
"Li et al., 2015 - DS",
"Teel et al., 2014 - Frontal - DS",
"Teel et al., 2014 - Parietal - DS",
"Teel et al., 2014 - Central - DS",
"Nwakamma et al., 2024 - Whole brain - DS",
"Nwakamma et al., 2024 - Fronto-parietal - DS",
"Lewine et al., 2019 - B",
"Franke et al., 2016 - Temporal - B",
"Franke et al., 2016 - Frontal - B",
"Modarres et al., 2016 - B",
"Alkhachroum et al., 2024 - NS",
"Arbour et al., 2015 - M")
```

## Number of participants in each frequency band

```{r}

# to help avoind double counting
moi_meta1 <- moi_meta |>
  mutate(study_base = str_remove(study_id, " -.*"))

meta_count <- moi_meta1 |> 
  distinct(study_base, beta_effect_size, tbi, controls) |>
  group_by(beta_effect_size) |>
  summarise(
    total_TBI = sum(tbi, na.rm = TRUE),
    total_control = sum(controls, na.rm = TRUE),
    total_N = total_TBI + total_control,
    k = n()  # number of studies/effect sizes
  )

meta_count

table(moi_meta1$study_base)
```

```{r}

#performed normality test and tested for significant differences in specific moderator variables to determine which variable to put into model to simple model output and interprebility.

shapiro.test(moi_meta$perc_tbi) # to detect normality of data

t.test(moi_meta$age_tbi[moi_meta$delta_effect_size], moi_meta$age_control[moi_meta$delta_effect_size]) # if data is parametric

# Or non-parametric if non-normal
wilcox.test(moi_meta$perc_tbi, moi_meta$perc_control, paired = TRUE)

# Used difference score in sex variable to input into model for simplicity
moi_meta$sex_diff <- moi_meta$perc_tbi - moi_meta$perc_control

moi_meta$severity <- as.factor(moi_meta$severity)

mlm_meta <- rma.mv(
  yi = theta_effect_size,
  V = theta_es_variance,
  mods = ~ age_tbi + sd_tbi,
  random = list( ~ 1 | study_id),  # Random effect for studies
  slab = study_labels,
  method = "REML",    # Restricted maximum likelihood estimation
  data = moi_meta)

# View results
summary(mlm_meta)

mlm_meta1 <- rma.mv(
  yi = delta_effect_size,
  V = delta_es_variance,
  random = list( ~ 1 | study_id),  # Random effect for studies
  slab = study_labels,
  method = "REML",    # Restricted maximum likelihood estimation
  data = moi_meta)

fit <- anova(mlm_meta1,mlm_meta, refit = TRUE)

fit

# View results
summary(mlm_meta1)


 # Forest plot
forest(mlm_meta)

```

```{r}

# Test for residual heterogeneity

mlm_meta1$slab <- study_labels

meta_eggers <- ranktest(mlm_meta1)

print(meta_eggers)
```

## Sample-based Funnel Plots with different p-value thresholds

```{r}

sei <- sqrt(moi_meta$alpha_es_variance[!is.na(moi_meta$alpha_es_variance) & !is.na(moi_meta$beta_es_variance)])

# Extract the data you need
effect_sizes <- moi_meta$alpha_effect_size[!is.na(moi_meta$alpha_effect_size)]
sample_sizes <- moi_meta$tbi[!is.na(moi_meta$alpha_effect_size)] + 
                moi_meta$controls[!is.na(moi_meta$alpha_effect_size)]  # Total N across both groups

max_n <- max(sample_sizes, na.rm = TRUE)
min_n <- min(sample_sizes, na.rm = TRUE)
n_sequence <- seq(min_n, max_n, length.out = 100)
precision <- sqrt(4/n_sequence)  # Approximation based on precision
mean_es <- mean(effect_sizes, na.rm = TRUE)

x_10 <- c(mean_es + 1.64 * precision, rev(mean_es - 1.64 * precision))
y_10 <- c(n_sequence, rev(n_sequence))

x_05 <- c(mean_es + 1.96 * precision, rev(mean_es - 1.96 * precision))
y_05 <- c(n_sequence, rev(n_sequence))

x_01 <- c(mean_es + 2.58 * precision, rev(mean_es - 2.58 * precision))
y_01 <- c(n_sequence, rev(n_sequence))

x_001 <- c(mean_es + 3.29 * precision, rev(mean_es - 3.29 * precision))
y_001 <- c(n_sequence, rev(n_sequence))



# Create overall sample-based funnel plot, which will be seperate from the frequency bands based ones.

moi_meta_longer <- moi_meta %>%
  pivot_longer(cols = -age_tbi, names_to = "Domains of bias", values_to = "Count")

bias_data$`Domains of bias` <- factor(bias_data$`Domains of bias`, 
                                      levels = c("Overall","Reported_result", "Outcome", "Missing_data","Post_interventions", 
                                                 "Selection","Exposure", "Confounding"))

bias_data$Risk_of_bias <- factor(bias_data$Risk_of_bias, 
                                 levels = c( "Very high risk", "High risk", "Some concerns", "Low risk"))


# Create the plot
plot(effect_sizes, sample_sizes,
     xlab = "Effect Size",
     ylab = "Sample Size",
     pch = 19,
     main = "Alpha Sample Size-Based Funnel Plot",
     ylim = c(min(sample_sizes) , max(sample_sizes)),
     xlim = c(min(effect_sizes) - 0.5, max(effect_sizes) + 0.5)) + 
abline(v = mean(effect_sizes, na.rm = TRUE), lty = "dashed")  + 
lines(mean_es - 1.64 * precision, n_sequence, col = "gray60",lwd = 2)+ 
lines(mean_es + 1.64 * precision, n_sequence, col = "gray60",lwd = 2)+
lines(mean_es + 1.96 * precision, n_sequence, col = "gray70", lwd = 2) + 
lines(mean_es - 1.96 * precision, n_sequence, col = "gray70",lwd = 2) +
lines(mean_es + 2.58 * precision, n_sequence, col = "gray80",lwd = 2) + 
lines(mean_es - 2.58 * precision, n_sequence, col = "gray80",lwd = 2) + 
lines(mean_es + 3.29 * precision, n_sequence, col = "gray90",lwd = 2) +
lines(mean_es - 3.29 * precision, n_sequence, col = "gray90",lwd = 2) +
  legend("topleft", 
       c("p < 0.10", "p < 0.05", "p < 0.01", "p < 0.001"),
       lty = c("solid", "solid", "solid", "solid"),
       col = c("gray90", "gray80", "gray70", "gray60"),
       lwd = 2) 

```
