---
title: "Mechanisms of Injury Meta-analysis"
format: html
editor: visual
---

## Using a meta-regression to determine whether TBI etiology produces distinct spectral power signatures 

```{r}
1 + 1
```

Necessary libraries

```{r}
library(metafor)
library(tidyverse)
library(readxl)
library(psych)
library(ggplot2)
library(dplyr)
library(sjPlot)
library(lmerTest)
library(lme4)
```

## Set directory and import data

Imported data frame detailing study, age, sex, severity, mechanism and all 5 bands effect sized and variance.

```{r}
setwd('/Users/nwakamma.m/Documents/manscripts/EEG & TBI review/meta_analysis')
moi_meta <- read_excel('eeg_tbi_meta_analysis.xlsx')
```

## Convert Cohen's D to Hedges G.

Very necessary to adjust for sample size when calculating effect size. Hedge's G takes sample size into account. The escalc package can get hedges G. Used variance estimation formula using hedge's g effect size.

```{r}
g <- esc::hedges_g(0.62, 63)   #input cohen's d and total sample size to get hedge's g

v <- (34 + 29)/(34 * 29) + g^2/(2*(34 + 29 - 2)) #get variance

g

v
```

## Multilevel Meta-regression with covariates

```{r}
moi_meta$mechanism <- as.factor(moi_meta$mechanism)
moi_meta$severity <- as.numeric(moi_meta$severity)
moi_meta$electrode_site <- as.numeric(moi_meta$electrode_site)

study_labels <- c("Thompson et al., 2005 - S",
"Balkan et al., 2015 - S",
"Munia et al., 2017 - S",
"Coyle et al., 2023 - NS",
"Fisher et al., 2015 - NS",
"Korn et al., 2005 - NS",
"Tebano et al., 1988 - NS",
"Williams et al., 2008 - M",
"Li et al., 2015 - DS",
"Teel et al., 2014 - Frontal - DS",
"Teel et al., 2014 - Parietal - DS",
"Teel et al., 2014 - Central - DS",
"Nwakamma et al., 2024 - Whole brain - DS",
"Nwakamma et al., 2024 - Fronto-parietal - DS",
"Lewine et al., 2019 - B",
"Franke et al., 2016 - Temporal - B",
"Franke et al., 2016 - Frontal - B",
"Modarres et al., 2016 - B",
"Alkhachroum et al., 2024 - NS",
"Arbour et al., 2015 - M")


mlm_meta <- rma.mv(
  yi = beta_effect_size,
  V = beta_es_variance,
  mods = ~ mechanism + age_tbi+ perc_tbi+ perc_control+time_since_injury,
  random = list( ~ 1 | study_id),  # Random effect for studies
  slab = study_labels,
  method = "REML",    # Restricted maximum likelihood estimation
  data = moi_meta)

# View results
summary(mlm_meta)


 # Forest plot
forest(mlm_meta)

# Test for residual heterogeneity
fit <- anova(mlm_meta,mlm_meta1, refit = TRUE)

fit

standard_mlm_meta <- rma(yi = alpha_effect_size, 
                     vi = alpha_es_variance,
                     slab = study_labels,
                     method = "REML", data = moi_meta)

summary(standard_mlm_meta)


standard_mlm_meta$slab <- study_labels

forest(standard_mlm_meta)

meta_eggers <- ranktest(standard_mlm_meta)

print(meta_eggers)
```
